\documentclass{article}

\usepackage{fullpage}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
%\usepackage[]{algorithm2e}
%\usepackage[linesnumbered]{algorithm2e}
\usepackage{enumerate,url}
\usepackage{hyperref}
\hypersetup{colorlinks=true}
\usepackage{enumitem}
\usepackage[ruled]{algorithm2e} % Added by Shahrokh for Pseudocodes
\usepackage{tcolorbox}


\title{CSE 6140 / CX 4140 Assignment 3\\due Oct 16, 2020 at 11:59pm on Canvas }
\author{Shasha Liao}
\date{}
\begin{document}
\maketitle


\section{Dominating set [12 pts]}
You’re configuring a large network of workstations, which we’ll model as
an undirected graph $G$; the nodes of $G$ represent individual workstations
and the edges represent direct communication links. The workstations all
need access to a common core database, which contains data necessary
for basic operating system functions.

You could replicate this database on each workstation; this would
make look-ups very fast from any workstation, but you’d have to manage
a huge number of copies. Alternately, you could keep a single copy of the
database on one workstation and have the remaining workstations issue
requests for data over the network $G$; but this could result in large delays
for a workstation that’s many hops away from the site of the database.

So you decide to look for the following compromise: You want to
maintain a small number of copies, but place them so that any workstation
either has a copy of the database or is connected by a direct link to a
workstation that has a copy of the database. In graph terminology, such
a set of locations is called a {\em dominating set}.

Thus we phrase the {\em Dominating Set Problem} as follows. Given the
network $G$, and a number $k$, is there a way to place $k$ copies of the database
at $k$ different nodes so that every node either has a copy of the database
or is connected by a direct link to a node that has a copy of the database?

Show that Dominating Set is NP-complete. Follow all steps we have outlined in class for a complete proof. \emph{Hint}: consider the Vertex Cover problem.

\begin{tcolorbox}
{\bf Solution:} 
\begin{itemize}
\item Step 1: Show that  {\em Dominating Set Problem} is in NP.\\
A potential solution would be $L_k = [v_1, v_2, ..., v_k]$, which is a list of $k$ vertices in the graph $G$ that was placed a copy of the database. To check if $L_k$ is a correct solution, we can loop through all the vertices in the $L_k$, store their neighbors in a hashset, and then check if the hashset has a length equal to $|V|$, the number of vertices in $G$. If we use a hashset to store $L_k$, then the worst runtime for checking a potential solution is $O(k|E|)$, where $E$ is the number of edges in $G$. Therefore, {\em Dominating Set Problem} is in NP. 

\item Step 2: Choose an NP-complete problem X.\\
{\em Vertex Cover: Given a graph $G = (V, E)$ and an integer $k$, dose there exist a subset of vertices $S \subseteq V$ with $|S| \leq k$ such that each edge in $E$ has at least one endpoint in $X$? }\\ 
We know the {\em Vertex Cover} problem is NP-complete. 
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}
\begin{itemize}
\item Step 3: Prove that  {\em Vertex Cover}  $ \leq_{p}$ {\em Dominating Set}.
\begin{itemize}
\item Given a {\em Vertex Cover} instance  $G = (V, E)$ and $k$, we construct a {\em Dominating Set} instance $G' = (V', E')$ and $k'$ that has a dominating set of size $k'$ iff $G$ has a vertex cover of size $k$. For each edge $e = (a,b)$ in $E$, it has at least one endpoint in $S$. We add a new vertex $v_{ab}$ between vertices $a$ and $b$ and connecting them with edges, i.e., we add two new edges $(a, v_{ab})$ and $(v_{ab}, b)$. In this way, we constructed our new graph $G' = (V', E')$. Also, note that if $G$ has isolated vertices $I = \{ v_i \in V | v_i \text{ is isolated in } G\}$, then these isolated vertices will not be included in a cover set of $G$ since they don't belong to any edge. However, a dominating set in $G$ will have to contain all the isolated vertices since there is no way for them to have a neighbor in the dominating set. Therefore, we need to set $k' = k + |I|$. Obviously, reducing an instance of {\em Vertex Cover} to an instance of {\em Dominating Set} only requires a time complexity of $O(|E|)$. Now we are ready to prove that $G = (V, E)$ has a cover set of size $k$ iff $G' = (V', E')$ has a dominating set of size $k'$.

\item "$\Rightarrow$" Let $X \subseteq V$ be a vertex cover of size $k$ in $G$. Then  $X \cup I $ is a dominating set of size $k'$ in $G'$. To show this, we prove by contradiction. Since $X \cup I $ has size $k'$ and every vertex in $I$ is for sure in the dominating set $X \cup I$. So the only way for $X \cup I $ not to be a dominating set in $G'$ is that there exists some vertex $u$ in $V' - I$ such that $u \notin X$ and all the neighbors of $u$ are also not in $X$. The way $G'$ was constructed ensures that $u$ has at least one neighbor $v \in V$. If  $u \in V$, then the edge $(u, v)$ is not covered by $X$, which contradicts with the fact that $X$ is a vertex cover in $G$. If $u \notin V$, then $u$ was added between two vertices $a, b \in V$, which means $u$ has only two neighbors $a$ and $b$ and neither $a$ nor $b$ are in $X$. Since there is an edge between $a$ and $b$ in $G$, we know this edge is not covered by $X$. Again, we get the contradiction. So $X \cup I $ is a dominating set of size $k'$ in $G'$. 

\item "$\Leftarrow$" Let  $X \cup I $ be a dominating set of size $k'$ in $G'$. If $X$ is not a vertex cover set of $G$, then there exists an edge $(a,b) \in E$ such that $a \notin X$ and $b \notin X$. Then the vertex $v_{ab}$ added between $a$ and $b$ dose not has a neighbor in $X \cup I$, which contradicts with the fact that $X \cup I$ is a dominating set in $G'$.
\end{itemize}
This completes the proof of {\em Vertex Cover}  $ \leq_{p}$ {\em Dominating Set}.
\end{itemize}
\end{tcolorbox}
\newpage
\section{Frenemies [12 pts]} 

Assume you are planning a dinner party and going to invite a set of friends. However, among them, there are some pairs of persons who are enemies. You need to create a seating plan and you are wondering if it is possible to arrange this set of $n$ friends of yours around a round table such that none of the two enemies will seat next to each other. Given the set of the $n$ friends and the set of the pairs of enemies, prove that this problem is NP-Complete. Remember to follow the steps from lecture to prove NP-completeness.

You can use the fact that \texttt{Hamiltonian Cycle (HC)} is NP-complete.

\begin{tcolorbox}
{\bf Solution:} 
\begin{itemize}
\item Step 1: Show that \texttt{Frenemies} is in NP.\\
A potential solution would be $L = \{a_1, a_2, ..., a_n\}$, which should be a permutation of numbers $1,2, ..., n$.  To check if $L$ is a correct solution, we can loop through $L$ and check if the numbers in $L$ are unique using a hashset and if $(a_i, a_{(i+1)\%n})$ is in the set of the pairs of enemies. This procedure takes $O(n)$ time.

\item Step 2: Choose an NP-complete problem: \texttt{Hamiltonian Cycle}.\\
\texttt{Hamiltonian Cycle}: Given an undirected graph $G = (V, E)$, does there exist a simple cycle that contains every node in $V$?

\item Step 3: Prove that \texttt{Hamiltonian Cycle} $\leq _{p}$ \texttt{Frenemies}.
\begin{itemize}
\item Given a \texttt{Hamiltonian Cycle} instance $G = (V, E)$, we construct a \texttt{Frenemies} instance. Suppose there are $n$ vertices in $G$, we consider $n$ friends in \texttt{Frenemies}. For each edge $(i,j)$ missing in $G$, i.e.,  $(i,j) \in E_c$, we construct a pair of enemies $(i, j)$ in \texttt{Frenemies}. Since there are $\frac {n(n-1)}{2}$ in a complete graph with $n$ vertices, the procedure of constructing all the pairs of enemies takes polynomial time. Now suppose $v_1, v_2, ..., v_n, v_1$ is a Hamiltonian Cycle of $G$, we can construct a solution to \texttt{Frenemies} be arranging the friends in the order of $v_1, v_2, ..., v_n, v_1$ around a table. This takes linear time. Now we are ready to prove that the two problems are equivalent. 

\item "$\Rightarrow$" Suppose $v_1, v_2, ..., v_n, v_1$ is a Hamiltonian Cycle of $G$, then arranging friends in the order of $v_1, v_2, ..., v_n, v_1$ around a table avoids two enemies sitting next to each other. Suppose not, say friends $v_{i}$ and $v_{i+1}$ are enemies but they sit next to each other. This is a contradiction because there will be no edge between $v_i$ and $v_{i+1}$ in $G$ if friends $v_i$ and $v_{i+1}$  are enemies. Therefore, arranging friends in the order of $v_1, v_2, ..., v_n, v_1$ gives a solution to \texttt{Frenemies}.

\item "$\Leftarrow$" Suppose arranging friends in the order of $v_1, v_2, ..., v_n, v_1$ is a solution to \texttt{Frenemies}. Since for sure $v_1, v_2, ..., v_n, v_1$ contains all the vertices in $G$, we just need to prove that it is a cycle in $G$.  If not, there exists an edge $(v_i, v_{i+1}) \notin E$, i.e.,  $(v_i, v_{i+1}) \in E_c$, which means friends $v_i$ and $v_{i+1}$ are enemies and they are sitting next to each other. This contradicts with the fact that arranging friends in the order of $v_1, v_2, ..., v_n, v_1$ is a solution to \texttt{Frenemies}. So $v_1, v_2, ..., v_n, v_1$ is a Hamiltonian cycle in $G$. 
\end{itemize} 
Therefore, we proved that \texttt{Frenemies} is NP-complete.
\end{itemize}
\end{tcolorbox}
\newpage
\section{Let's go hiking [26 pts]}
Alex and Baine go hiking together. They bring a bag of items and want to divide them up. For the following scenarios, decide whether the problem can be solved in polynomial time. If yes, give a polynomial-time algorithm; otherwise prove the problem is NP-complete. 
\begin{itemize}
	\item (8 pts) The bag contains $n$ items of two weights: 1lb and 2lb. Alex and Baine want to divide the items evenly so that they carry the same amount of weight. 
	
\begin{tcolorbox}
{\bf Solution:} This problem can be solved in constant time. Suppose there are $n_1$ items of 1lb and $n_2$ items of 2lb with $n_1 + n_2 = n$. There are three different cases: 
\begin{itemize}
\item Case 1: $n_1$ is odd. \\
There is no way for Alex and Baine to divide the items evenly since the total number of weights is odd.
\item Case 2: $n_1$ and $n_2$ are both even. \\
Alex and Baine can divide the items evenly simply by each of them taking $n_1/2$ items of 1lb and $n_2/2$ items of 2lb.
\item Case 3: $n_1$ is even but $n_2$ is odd. \\
If $n_1 = 0$, there is no way to divide the items evenly. Otherwise, they can divide the items evenly by treating 2 items of 1lb as 1 item of 2lb and reduce the problem to Case 2.
\end{itemize}
\end{tcolorbox}
\newpage	
	\item (9 pts) The bag contains $n$ items of different weights. Again they want to divide the items evenly. 
\begin{tcolorbox}
{\bf Solution:} Denoting this problem by $Y$, we can prove that it is NP-complete by showing $\texttt{Subset Sum} \leq_p  Y$.
\begin{itemize}
\item Step 1: Prove that $Y$ is in NP. \\
Suppose the bag contains $n$ items of weights $\{w_1, w_2, ..., w_n\}$. A potential solution would be that Alex carries a subset of items with weights $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ where $0 < p < n$, and Baine carries the rest of the items. To check if it is a correct solution, we only need to check if $\sum_{i=1}^p w_{k_i} = \frac {\sum_{i=1}^n w_i }{2}$, which takes $O(n)$ time.
\item Step2: Choose an NP-complete problem:  \texttt{Subset Sum}. \\
 \texttt{Subset Sum}: given natural numbers $w_1, w_2, ..., w_n$ and a target $W$, dose there exist a subset of $\{w_1, w_2, ..., w_n\}$ that adds up to exactly $W$?
 
\item Step 3: Prove that $\texttt{Subset Sum} \leq_p  Y$.
\begin{itemize}
\item Reduction: given an instance $(w_1, w_2, ..., w_n, W)$ of \texttt{Subset Sum}, we can create an instance of $Y$ as this: the bag contains $n+1$ items of weights $\{w_1, w_2, ..., w_n, w_{n+1}\}$ with $$w_{n+1} = \sum_{i=1}^n w_i - 2W.$$ We will show that if the  \texttt{Subset Sum} problem has a solution $\{w_{k_1}, w_{k_2}, ..., w_{k_p}\}$, then we can divide $n+1$ items of weights $\{w_1, w_2, ..., w_n, w_{n+1}\}$ evenly by taking one part as $\{w_{k_1}, w_{k_2}, ..., w_{k_p}, w_{n+1}\}$, and the rest as the other part.
\item "$\Rightarrow$" If $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ is a solution to the \texttt{Subset Sum} problem, i.e., $\sum_{i=1}^p w_{k_i} = W$. Then we have $$\sum_{i=1}^p w_{k_i} + w_{n+1} = W + (\sum_{i=1}^n w_i - 2W) = \sum_{i=1}^n w_i - W = \frac{\sum_{i=1}^n w_i + w_{n+1}}{2}.$$ So $\{w_{k_1}, w_{k_2}, ..., w_{k_p}, w_{n+1}\}$ is a solution to our problem $Y$. 
\item "$\Leftarrow$" If $\{w_{k_1}, w_{k_2}, ..., w_{k_p}, w_{n+1}\}$ is a solution to our problem $Y$, then $$\sum_{i=1}^p w_{k_i} + w_{n+1} =  \frac{\sum_{i=1}^n w_i + w_{n+1}}{2}.$$ So $$\sum_{i=1}^p w_{k_i} =  \frac{\sum_{i=1}^n w_i + w_{n+1}}{2} - w_{n+1} =   \frac{\sum_{i=1}^n w_i - w_{n+1}}{2} = W.$$ This proved that $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ is a solution to the \texttt{Subset Sum} problem.
\end{itemize}
\end{itemize}
Therefore, $\texttt{Subset Sum} \leq_p  Y$ and $Y$ is NP-complete.
\end{tcolorbox}
\newpage
	\item (9 pts) The bag contains $n$ items of different weights. They want to divide the items such that the weight difference of items they carry is less than 10lbs. 
\begin{tcolorbox}
{\bf Solution:} Denoting this problem by $Z$, we can prove that it is NP-complete by showing $\texttt{Subset Sum} \leq_p  Z$.
\begin{itemize}
\item Step 1: Prove that $Z$ is in NP. \\
Suppose the bag contains $n$ items of weights $\{w_1, w_2, ..., w_n\}$. A potential solution would be that Alex carries a subset of items with weights $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ where $0 < p < n$, and Baine carries the rest of the items. To check if it is a correct solution, we only need to check if $\frac {M- 10 }{2} < \sum_{i=1}^p w_{k_i} < \frac {M+ 10 }{2}$ with $M = \sum_{i=1}^n w_i $, which takes $O(n)$ time.
\item Step2: Choose an NP-complete problem:  problem $Y$ in the previous part. \\
 $Y$: The bag contains $n$ items of different weights with each weight being a natural number. Alex and Baine want to divide the items evenly.  
\item Step 3: Prove that $Y \leq_p  Z$.
\begin{itemize}
\item Reduction: given an instance $\{w_1, w_2, ..., w_n\}$ of $Y$, we can create an instance of $Z$ as this: the bag contains $n$ items of weights $\{w'_1, w'_2, ..., w'_n\}$ with $$w'_{i} = 11 \times w_i.$$ We will show that if problem $Y$ has a solution $\{w_{k_1}, w_{k_2}, ..., w_{k_p}\}$ iff problem $Z$ has a solution $\{w'_{k_1}, w'_{k_2}, ..., w'_{k_p}\}$.

\item "$\Rightarrow$" If $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ is a solution to problem $Y$, i.e., $\sum_{i=1}^p w_{k_i} = \frac {M }{2}$ with $M = \sum_{i=1}^n w_i $. Then we have $$\sum_{i=1}^p w'_{k_i} = \sum_{i=1}^p 11w_{k_i}= \frac{11M}{2} = \frac{\sum_{i=1}^n 11w_i}{2} = \frac{M'}{2}$$ with $M' = \sum_{i=1}^n w'_i = 11M$. Since $\sum_{i=1}^p w'_{k_i} = \frac{M'}{2} \in ( \frac {M' - 10 }{2}, \frac {M'+ 10 }{2} )$, we know $\{w'_{k_1}, w'_{k_2}, ..., w'_{k_p}\}$ is a solution to our problem $Z$. 

\item "$\Leftarrow$" If $\{w'_{k_1}, w'_{k_2}, ..., w'_{k_p}\}$ is a solution to our problem $Z$, then $$\sum_{i=1}^p w_{k_i}=\frac{1}{11}  \sum_{i=1}^p w'_{k_i} \in (  \frac {M' - 10 }{22}, \frac {M'+ 10 }{22} ) = (\frac{M}{2} - \frac{10}{22}, \frac{M}{2} + \frac{10}{22}).$$ So $$\sum_{i=1}^p w_{k_i} =  \frac{M}{2}.$$ This proved that $\{w_{k_1}, w_{k_2}, ..., w_{k_p} \}$ is a solution to problem $Y$. 
\end{itemize}
\end{itemize}
Therefore, $Y \leq_p  Z$ and $Z$ is NP-complete.
\end{tcolorbox}
\end{itemize}


\textbf{Hint}: Recall Subset Sum problem: given a set $X$ of integers and a target number $t$, find a subset $Y\subset X$ such that the members of $Y$ add up to exactly $t$. 

\end{document}
